\batcave: Calling somatic mutations with a tumor- and site-specific prior

### 2019-12-16 ResponseToReviews.tex

Lastly, as correctly pointed out by the second reviewer, the sequencing coverage only makes sense in the context of the purity of the samples and this information should be provided to the reader.
\begin{response}
\RNG{Brian: Can you think about this in the context of what MuTect already does?}
\BKM{Mutect has a parameter for contamination/purity if known. I could add a parameter default=0, and write a message to stdout with the value used. There are various methods for estimating purity.}
\end{response}

\section*{Reviewer 1}

Mannakee and Gutenkunst present BATCAVE, an algorithm that first estimates the tumor’s mutation profile and mutation rate using high-confidence variants and then uses them as a prior when calling other variants based on mutational signatures. The concept behind this work is interesting. However, the difference in accuracy between MuTect and BATCAVE is relatively small except for the case when one has particularly high depth sequencing (500x), limiting the utility of the method.

Still, the 500x whole exome case does indicate that BATCAVE is better at identifying low frequency variants and thus can better reveal intratumoral heterogeneity, an important area of inquiry in cancer. Interestingly, the AUROC values for both MuTect and BATCAVE have room for improvement for the 500x whole exome case (Table 1). I suggest the authors further investigate how to improve the AUROC values. For example, in equation 4, it is unclear why the authors assume $\nu$ be uniformly distributed. Although this is also an assumption of MuTect, the BATCAVE approach inherently distinguishes mutations by allele frequency – since log-likelihood is closely associated to AF.  Perhaps a model in which $\nu$ follows a decreasing distribution vs. AF (e.g. in concordance with a model of a growing tumor) would yield better AUROC values.  
\begin{response}
A uniform prior over $\nu$ is appropriate when the goal is to make the best possible call at each potential variant site.
Conceptually, before examining the read data we have no information about the allele frequency at any individual site.
Practically, a nonuniform prior would favor low-frequency and thus typically low-evidence sites, increasing the false positive rate.
If our goal were to best estimate the distribution of variant allele frequencies in the tumor, then a nonuniform $\nu$ would indeed be important.

We discuss this issue in a new paragraph in Discussion.
\end{response}

In Rubanova et al ( https://www.biorxiv.org/content/10.1101/260471v4.full), it is argued that there are changes in mutational signatures over time in cancers. While Mannakee and Gutenkunst mention this work, they should further comment on whether such changes in mutational signatures over time are small enough to be explained by the BATCAVER concept of re-evaluating mutations based on p-value.
\begin{response}
\RNG{Going to be difficult, depending on biology to offer specific guidance about the size of the change. Rather, we now do a goodness-of-fit test and warn the user if it looks like it's violated.}
\BKM{We could detect a change above. Whether it matters or not is a complex interplay between depth and selection. For late changes most mutations with that signature will be below the threshold of detection. For an early mutation all of the marginal (low frequency) variants will be from the signature.}
\end{response}

Minor
p.2: “Folding the central base to the pyrimidines”: The “folding” terminology is not standard and should be improved. 
\begin{response}
Corrected
\end{response}

### 2019-12-16 NAR_DRAFT.tex
Response to uniform nu comment

\rngcomment{Assumption of uniform prior with $\nu$. Standard (cite MuTect). Also distinguish between calling individual mutations and inferring distribution of allele frequencies, like ANGSD.}

@@

>>>>

@@

\rngcomment{
Following MuTect \cite{Cibulskis2013}, \batcave assumes a uniform prior on the variant allele frequency.
In general, low-frequency variants are expected to be more common than high-frequency variants.
Nevertheless, the uniform frequency prior is appropriate when the goal is to make the best call possible at each potential variant site.
By favoring low-frequency and thus typically low-evidence variants, a nonuniform prior would increase the false positive rate.
Note however, that if the goal were not to best identify individual variants, but rather to best estimate the distribution of variant allele frequencies within the tumor, then a nonuniform frequency prior would be appropriate \cite{Nielsen2012}.
}
